---
title: "Initial Korean covid 19"
author: "Tyler Frasier"
date: "fall 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
 
```{r}
str(kRdata)
```
 
 
## Your Question - Is dying independent of sex?

$H_0$ 
there is no difference in dying from covid 19 based on gender.

$A_0$
there is a difference in dying from covid 19 based on gender.

## Methhods 


since sex and the response is cat~cat therefore we use an analysis called cat~cat
## Graphical results - use ggplot commands to do this  

```{r}
# Load the ggplot2 package
library(ggplot2)

# Create a grouped bar chart
ggplot(kRdata, aes(x = result, fill = Sex)) +
  geom_bar(position = "dodge") +
  labs(x = "Result", y = "Sex") +
  ggtitle("Grouped Bar Chart of Sex and Result")


```






```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")


basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

this plot shows more men die from covid 19 than woman.
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

In the table twice as many males ahd died than woman.


## Inferential results 
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

if you are male you are 2.3 timnes more likely to die from covid 19 than if you were female.


## Conclusions

is dying independent of sex well more men died from covid 19 inb korea than females so therefore men who travel to korea should be careful